\documentclass[11pt,a4paper]{article}

% ---- packages -----
\usepackage[a4paper,margin=2.5cm]{geometry}
\usepackage[onehalfspacing]{setspace}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{booktabs}
\usepackage{tabularray}
\UseTblrLibrary{booktabs}
\UseTblrLibrary{siunitx}
\usepackage{float}
\usepackage{graphicx}
\usepackage{amsmath, amssymb}
\usepackage{natbib}
\usepackage[hidelinks]{hyperref}
\usepackage{titlesec}
\usepackage{tikz}
\usetikzlibrary{positioning, arrows.meta}
\titlespacing*{\section}{0pt}{1.0\baselineskip}{0.4\baselineskip}
\titlespacing*{\subsection}{0pt}{0.8\baselineskip}{0.3\baselineskip}

% modelsummary helpers
\providecommand{\tinytableTabularrayUnderline}[1]{\underline{#1}}
\providecommand{\tinytableTabularrayStrikeout}[1]{\sout{#1}}

\title{\textbf{MSS (2004) Revisited:\\
       Weak Instruments and Dynamic Panel Identification}}
\author{Alexander Lievens and Tobias Vanoverschelde\\
\small KU Leuven: D0S91a Advanced Applied Econometrics\\
\small Prof.\ Dr.\ Dirk Czarnitzki}
\date{May 2026}

\begin{document}
\maketitle

\begin{abstract}
\noindent
We re-examine the rainfall-instrumental-variables design of
\citet{MSS2004} using post-2005 weak-instrument diagnostics and an
alternative panel identification strategy taught in our course. On the
original panel of 41 sub-Saharan African countries (1981--1999), the
rainfall first stage fails the \citet{StockYogo2005} 10\%-maximum-size
critical value by a factor of seven; the 2SLS point estimate carries
the wrong sign and a cluster-bootstrap percentile interval that
covers the entire interesting parameter space. A first-difference IV
with the lagged growth level \citep{AndersonHsiao1981}, which does not
depend on rainfall, recovers the negative MSS sign ($\hat\beta_{AH} =
-0.35$) but with comparable imprecision. The data inform the sign of
the income-conflict relationship but not its magnitude under either
identifying assumption available within the original design.
\end{abstract}

% ===========================================================================
\section{Introduction}\label{sec:intro}
% ===========================================================================
Estimating the causal effect of income shocks on civil conflict is
not an estimation problem but an identification problem. OLS confounds
three biases: conflict destroys output (reverse causation), national
accounts in low-income economies are measured with error, and
persistent country traits affect both (omitted variables). No amount
of data fixes this. The effect is identified only if we can find a
source of variation in growth that is exogenous to conflict and does
not enter the conflict equation directly.

\citet{MSS2004}, henceforth MSS, proposed that source: annual rainfall
growth, used as an excluded instrument for per-capita GDP growth on a
panel of 41 sub-Saharan African countries over 1981--1999. The
identifying assumptions are that rainfall shifts agricultural output
(relevance), and that it affects conflict only through that channel
(exclusion). The paper was novel in 2004. It also predates several of
the diagnostic tools we now use to evaluate IV designs.

The literature that followed targeted each of these assumptions in
turn. \citet{BrucknerCiccone2010} proposed an alternative instrument
based on global commodity prices, relying on a different exogeneity
argument. \citet{Ciccone2011} questioned the specification of the
first stage, showing that rainfall's mean-reversion contaminates the
growth-as-IV interpretation. \citet{Sarsons2015} tested the exclusion
restriction directly, exploiting irrigation infrastructure that breaks
the agricultural channel.

We revisit MSS with two questions. First, how much the original
rainfall design tells us about $\beta$ once weak-instrument diagnostics
and a cluster bootstrap are applied to the standard 2SLS output.
Second, whether a first-difference IV strategy with lagged levels
\citep{AndersonHsiao1981}, which does not rely on rainfall, recovers
a coherent estimate.

Section~\ref{sec:identification} formalises the MSS identification.
Section~\ref{sec:critiques} surveys the critiques.
Section~\ref{sec:data} describes the data. Section~\ref{sec:methods}
sets out the estimators. Section~\ref{sec:results} reports estimates.
Section~\ref{sec:discussion} concludes.

% ===========================================================================
\section{The MSS identification strategy}\label{sec:identification}
% ===========================================================================
MSS fit the structural equation
\begin{equation}
C_{it} = \alpha_i + \lambda_i t + \beta\, g^{y}_{it} + \gamma' X_{it} + u_{it},
\label{eq:structural}
\end{equation}
where $C_{it} = \mathbf{1}\{\text{conflict}_{it}\}$ is a binary indicator for civil
conflict in country $i$ at year $t$, $g^{y}_{it} = \Delta \log(\text{GDP}_{it})$ is
per-capita GDP growth, $\alpha_i$ are country fixed effects, $\lambda_i t$ are
country-specific linear time trends, and $X_{it}$ collects time-varying controls
(lagged Polity~2). Opportunity-cost theory \citep{Grossman1991} predicts
$\beta < 0$.

Estimating equation~\eqref{eq:structural} by OLS is inconsistent for the reasons
listed in Section~\ref{sec:intro}. MSS instrument $g^{y}_{it}$ with annual
rainfall growth $g^{R}_{it}$ and its first lag:
\begin{equation}
g^{y}_{it} = a_i + b_i t + \pi_1\, g^{R}_{it} + \pi_2\, g^{R}_{i,t-1}
             + \delta' X_{it} + v_{it}.
\label{eq:firststage}
\end{equation}

The identifying assumptions, made conditional on $\alpha_i$ and $X_{it}$, are:

\textit{Relevance}: $\mathrm{Cov}(g^{R}_{it}, g^{y}_{it}) \neq 0$. Rainfall shifts
agricultural output, which is a large share of African GDP. Testable through the
first-stage $F$ statistic on the excluded instruments.

\textit{Exclusion}: $\mathrm{Cov}(g^{R}_{it}, u_{it}) = 0$. Rainfall affects
conflict only through GDP. Not directly testable in general; with two
instruments, the Sargan test provides one degree of freedom of indirect check on
whether the two rainfall measures imply the same $\beta$.

\begin{figure}[H]
\centering
\begin{tikzpicture}[
  node distance=2.6cm,
  every node/.style={font=\small},
  arr/.style={->, >={Stealth[length=2.5mm]}, thick}
]
  \node (z)  {$z$};
  \node[right of=z]  (y2) {$y_2$};
  \node[right of=y2] (y1) {$y_1$};
  \node[below=1.0cm of y2, xshift=1.3cm] (u)  {$u$};

  \draw[arr] (z)  -- (y2);
  \draw[arr] (y2) -- (y1);
  \draw[arr, dashed] (u) -- (y2);
  \draw[arr, dashed] (u) -- (y1);

  \node[above=0.05cm of z]  {\scriptsize rainfall};
  \node[above=0.05cm of y2] {\scriptsize growth};
  \node[above=0.05cm of y1] {\scriptsize conflict};
  \node[right=0.1cm of u]   {\scriptsize unobserved};
\end{tikzpicture}
\caption{The MSS identification strategy as a DAG. Solid arrows show the
assumed causal pathway: rainfall ($z$) affects growth ($y_2$), which affects
conflict ($y_1$). Dashed arrows show unobserved confounders ($u$) that bias
OLS. Identification of $\beta$ requires the $z \to y_2$ arrow (relevance,
testable) and the \emph{absence} of any $z \to y_1$ arrow (exclusion, not
directly testable).}
\label{fig:dag-mss}
\end{figure}

The specification choices matter. MSS use country fixed effects to absorb
time-invariant heterogeneity (geography, institutions, colonial history) and
country-specific linear time trends to absorb slow secular movements
(democratisation, technology diffusion). They do not use year fixed effects,
which would absorb the common component of rainfall shocks across the panel
and reduce the variation the instrument exploits. Standard errors are clustered
at the country level.

Their preferred 2SLS specification yields $\hat\beta \approx -2.4$: a
five-percentage-point decline in growth raises the probability of conflict by
roughly twelve percentage points.

% ===========================================================================
\section{Methodological critiques and alternative identification}\label{sec:critiques}
% ===========================================================================
The literature that followed MSS extended the design along three axes:
it offered alternative instruments, questioned the first-stage
specification, and tested the exclusion restriction directly. Only the
last two are critiques in the strict sense; the first is a parallel
identification strategy.

\textbf{Alternative instrument.} \citet{BrucknerCiccone2010} replace
rainfall with country-specific commodity export price shocks: the
weighted sum of global commodity prices, using each country's export
composition as weights. The exogeneity argument is small-country
price-taking. Global commodity prices are determined on world markets,
and no single sub-Saharan African country is large enough to move them.
Their first stage becomes
\begin{equation}
g^{y}_{it} = a_i + \tilde\pi\, \text{shock}^{P}_{it}
             + \tilde\delta' X_{it} + v_{it},
\label{eq:bc-fs}
\end{equation}
while the structural equation~\eqref{eq:structural} is unchanged. They
recover a negative and significant $\beta$, corroborating the MSS sign
under a different identifying assumption. This is identification
pluralism rather than critique: two unrelated exogeneity arguments
converge on the same conclusion, which strengthens MSS rather than
weakens it. Within MSS's own design, the analogue is the Sargan
over-identification test on the two rainfall instruments
(current and lagged); we return to it in Section~\ref{sec:results}.

\textbf{Specification critique.} \citet{Ciccone2011} attacks the
functional form rather than the exogeneity. He shows that the original
MSS result stems from a counterintuitive \emph{positive} correlation
between conflict at $t$ and rainfall levels at $t-2$, rather than from
negative rainfall shocks lowering income and raising conflict through
the agricultural channel. Re-examining the analysis with updated data,
he reports that conflict is largely unrelated to rainfall.
\citet{MiguelSatyanath2011} reply showing that the MSS findings hold
when rainfall levels are used as instruments, and that the first-stage
relationship weakens after 2000. The lesson is generic: an exogenous
instrument applied in the wrong functional form still produces an
inconsistent estimate.

\textbf{Direct test of exclusion.} The most consequential critique
tests the exclusion restriction itself. \citet{Sarsons2015} uses
Indian district-level data on rainfall and Hindu-Muslim religious
riots, combined with records of irrigation dams. In districts
\emph{downstream of dams}, water supply is decoupled from local
rainfall: income in these districts is much less sensitive to rainfall
fluctuations, so the agricultural channel is mechanically attenuated.
If MSS's exclusion holds (rainfall affects conflict only through
agricultural income), then in dam-downstream districts rainfall should
not predict conflict. She finds that rain shocks remain equally strong
predictors of riot incidence in downstream districts. The implication is a
direct $z \to y_1$ arrow that the MSS DAG denies (Figure~\ref{fig:dag-sarsons}).

\begin{figure}[H]
\centering
\begin{tikzpicture}[
  node distance=2.6cm,
  every node/.style={font=\small},
  arr/.style={->, >={Stealth[length=2.5mm]}, thick}
]
  \node (z)  {$z$};
  \node[right of=z]  (y2) {$y_2$};
  \node[right of=y2] (y1) {$y_1$};
  \node[below=1.0cm of y2, xshift=1.3cm] (u)  {$u$};

  \draw[arr] (z)  -- (y2);
  \draw[arr] (y2) -- (y1);
  \draw[arr, dashed] (u) -- (y2);
  \draw[arr, dashed] (u) -- (y1);
  \draw[arr, thick] (z) to[bend left=35] (y1);

  \node[above=0.05cm of z]  {\scriptsize rainfall};
  \node[above=0.05cm of y2] {\scriptsize growth};
  \node[above=0.05cm of y1] {\scriptsize conflict};
  \node[right=0.1cm of u]   {\scriptsize unobserved};
\end{tikzpicture}
\caption{Sarsons's critique as a DAG. Same diagram as
Figure~\ref{fig:dag-mss}, plus a direct $z \to y_1$ arrow that exclusion
denies. Evidence from regions where the $z \to y_2$ channel is
physically blocked (downstream of dams) supports the existence of this
arrow.}
\label{fig:dag-sarsons}
\end{figure}

The methodological lesson is general. Exclusion restrictions are
sometimes partially testable: if the IV is claimed to operate through
one channel, find a setting where the channel is shut and check whether
the IV still moves the outcome. The post-2015 standard for IV papers in
development economics has internalised this test.

% ===========================================================================
\section{Data}\label{sec:data}
% ===========================================================================
We use the replication panel distributed by MSS via the Harvard
Dataverse (doi:10.7910/DVN/27324). It covers 41 sub-Saharan African
countries over 1981--1999. After constructing one lag of GDP growth and
dropping country-years with missing controls, the estimation sample is
702 country-year observations.

The outcome \texttt{any\_prio} is a PRIO indicator equal to one if a
country experiences at least 25 battle deaths in a year. The endogenous
regressor \texttt{gdp\_g} is annual per-capita GDP growth. The excluded
instruments are current and lagged rainfall growth, \texttt{GPCP\_g}
and \texttt{GPCP\_g\_l}, constructed from the Global Precipitation
Climatology Project. Controls follow MSS: initial log income, lagged
Polity~2, ethnic and religious fractionalisation, an oil-exporter
dummy, log mountainous terrain, and log lagged population. Only lagged
Polity~2 is time-varying; the rest are absorbed by country fixed
effects in our preferred specifications.

\begin{table}[H]
\centering
\caption{Descriptive statistics, estimation sample ($N = 702$).}
\label{tab:descriptives}
\begin{tabular}{lrrrrr}
\toprule
Variable & $N$ & Mean & SD & Min & Max \\
\midrule
\texttt{any\_prio}  & 702 & 0.269   & 0.444 & 0.000   & 1.000 \\
\texttt{gdp\_g}     & 702 & $-0.006$ & 0.066 & $-0.474$ & 0.319 \\
\texttt{GPCP\_g}    & 702 & 0.015   & 0.210 & $-0.550$ & 1.680 \\
\texttt{y\_0}       & 702 & 1.16    & 0.90  & 0.316   & 4.84  \\
\texttt{polity2l}   & 702 & $-3.55$ & 5.55  & $-10$   & 9     \\
\texttt{ethfrac}    & 702 & 0.655   & 0.237 & 0.036   & 0.925 \\
\texttt{relfrac}    & 702 & 0.487   & 0.185 & 0.000   & 0.783 \\
\texttt{Oil}        & 702 & 0.117   & 0.321 & 0       & 1     \\
\texttt{lmtnest}    & 702 & 1.58    & 1.43  & 0       & 4.42  \\
\texttt{lpopl1}     & 702 & 8.77    & 1.20  & 5.75    & 11.7  \\
\bottomrule
\end{tabular}
\end{table}

Three features of the panel matter for what follows. The unconditional
conflict frequency is 27\%. Average growth is slightly negative
($-0.6\%$ per annum). The panel is democracy-poor on average (mean
Polity~2 of $-3.5$ on a scale of $-10$ to $10$).

% ===========================================================================
\section{Methodology}\label{sec:methods}
% ===========================================================================
Each estimator below corresponds to a different way of reading the DAG
in Figure~\ref{fig:dag-mss}. POLS uses all variation jointly and
trusts that the dashed $u \to y_2$ and $u \to y_1$ arrows are weak.
FE removes the time-invariant part of $u$. 2SLS exploits the $z \to
y_2$ arrow and assumes no direct $z \to y_1$ arrow. The cluster
bootstrap provides an alternative inference for $\hat\beta_{\text{2SLS}}$
when conventional asymptotics are unreliable. First-difference IV with
lagged levels identifies $\beta$ without using $z$ at all, relying on
internal panel variation instead.

\subsection{Pooled OLS (LPM)}\label{sec:pols}

We begin by estimating equation~\eqref{eq:structural} by OLS, treating
the binary outcome as a linear probability model. In subscript form,
\[
\hat\beta_{\text{POLS}}
   = \Big( \textstyle\sum_{it} x_{it} x_{it}' \Big)^{-1}
     \textstyle\sum_{it} x_{it} C_{it}
\quad\Longleftrightarrow\quad
\hat\beta_{\text{POLS}} = (X'X)^{-1} X' C,
\]
where the right-hand expression stacks all $NT$ observations into
matrices $X$ and $C$. Consistency requires
$\mathrm{Cov}(g^{y}_{it}, u_{it}) = 0$, which the three endogeneity
sources in Section~\ref{sec:intro} rule out. POLS
therefore serves as a benchmark, not as a causal estimate. Two issues
are handled at the inference stage. First, the LPM is heteroscedastic
by construction since $\mathrm{Var}(u \mid x) = p(x)(1-p(x))$ depends
on $x$; we report the Breusch-Pagan and White tests as confirmation.
Second, errors are correlated within country across years; we cluster
standard errors at the country level throughout.

\subsection{Fixed and random effects}\label{sec:fe}

Fixed-effects estimation removes the time-invariant part of $u$. Let
$u_{it} = c_i + e_{it}$ with $c_i$ constant within country. The within
transformation subtracts country means from each variable: let
$\tilde y_{it} = y_{it} - \bar y_i$ and $\tilde x_{it} = x_{it} - \bar
x_i$. The resulting estimator is
\[
\hat\beta_{\text{FE}}
   = \Big( \textstyle\sum_{it} \tilde x_{it} \tilde x_{it}' \Big)^{-1}
     \textstyle\sum_{it} \tilde x_{it} \tilde y_{it},
\]
which is consistent if $\mathrm{Cov}(g^{y}_{it}, e_{it}) = 0$ after
demeaning. We report two-way FE (country and year) as our baseline and
add a specification with country-specific linear trends in place of
year FE, the original MSS choice. The two differ in how they treat the
common component of rainfall variation across the panel: year FE
absorb it, country trends preserve it.

Random effects is consistent under the stronger assumption
$\mathrm{Cov}(c_i, x_{it}) = 0$ and is more efficient when it holds.
The Hausman test compares FE and RE: a large $\chi^2$ statistic
indicates the RE assumption fails. Time-invariant controls (initial
income, fractionalisation, terrain, oil, lagged population) are
absorbed by country FE and do not appear in the FE or RE
specifications.

\subsection{Two-stage least squares}\label{sec:2sls}

We then implement the MSS strategy: instrument $g^{y}_{it}$ with
current and lagged rainfall growth from equation~\eqref{eq:firststage}.
As a two-stage procedure, run the first stage in
equation~\eqref{eq:firststage} by OLS to obtain fitted values
$\hat g^{y}_{it}$, then plug those into the structural equation:
\[
C_{it} = \alpha_i + \lambda_i t + \beta\, \hat g^{y}_{it}
       + \gamma' X_{it} + \text{error}_{it}.
\]
In matrix form this is equivalent to
\[
\hat\beta_{\text{2SLS}} = (X' P_Z X)^{-1} X' P_Z C,
\]
where $P_Z = Z(Z'Z)^{-1} Z'$ projects onto the column space of the
instruments (country FE and exogenous controls included in $Z$). Two
assumptions support consistency.

\textit{Relevance.} We test it through the cluster-robust $F$ statistic
on the excluded instruments in the first stage \citep{StaigerStock1997},
compared to the \citet{StockYogo2005} 10\%-maximum-size critical value
of 19.93 (two IVs, one endogenous regressor). Values below the
threshold indicate weak instruments and invalid conventional standard
errors.

\textit{Exclusion.} Not directly testable. With two instruments we
report the Sargan over-identification test, which asks whether both
rainfall measures imply the same $\beta$. Rejection indicates that at
least one instrument is correlated with $u$. Failure to reject is weak
evidence at best given the single degree of freedom and the fact that
both instruments share the same exogeneity argument; the test cannot
detect homogeneous invalidity.

We additionally report the Wu-Hausman endogeneity test, which compares
OLS and 2SLS and rejects exogeneity of $g^{y}_{it}$ if the two
estimates differ by more than sampling noise.

\subsection{Cluster bootstrap inference}\label{sec:boot}

When the first-stage $F$ falls short of conventional critical values,
the standard 2SLS $t$-statistic does not have an asymptotic standard
normal distribution and reported confidence intervals can have actual
coverage far below their nominal level. We supplement the conventional
output with a non-parametric cluster bootstrap. Countries are resampled
with replacement, the 2SLS coefficient is re-estimated on each
resampled panel, and the percentile interval is reported from $B=500$
replications. Resampling at the country level preserves the
within-country correlation structure that the cluster-robust standard
errors of Section~\ref{sec:pols} assume.

\subsection{First-difference IV with lagged levels}\label{sec:ahiv}

If rainfall is too weak to identify $\beta$ externally, the panel
itself contains a second source of variation. First-differencing
equation~\eqref{eq:structural} eliminates the country fixed effects:
\[
\Delta C_{it} = \beta\, \Delta g^{y}_{it} + \gamma' \Delta X_{it}
              + \Delta u_{it}.
\]
\citet{AndersonHsiao1981} propose instrumenting the differenced
regressor $\Delta g^{y}_{it}$ with its lagged level $g^{y}_{i,t-1}$,
under the assumption $\mathrm{Cov}(g^{y}_{i,t-1},\Delta u_{it}) = 0$:
past growth is uncorrelated with the change in unobserved conflict
shocks. This identifying assumption is logically separate from the
rainfall exclusion restriction; the two strategies fail or succeed for
different reasons. The Sargan and Wu-Hausman diagnostics from
Section~\ref{sec:2sls} apply unchanged.

% ===========================================================================
\section{Results and diagnostics}\label{sec:results}
% ===========================================================================
Table~\ref{tab:main} reports the seven static specifications described
in Section~\ref{sec:methods}. Tables~\ref{tab:diag}
and~\ref{tab:gmm} report the diagnostic battery and the dynamic-panel
GMM estimates.

\begin{table}[H]
\centering
\caption{Static specifications. Cluster-robust SE at country level in
parentheses. Models 2--6 use two-way FE; model~7 replaces year FE with
country-specific linear time trends. Time-invariant controls are
included in POLS only and absorbed by country FE elsewhere.}
\label{tab:main}
\small
\begin{tabular}{lrrrrrrr}
\toprule
 & POLS & FE & RE & Red.\ form & First stage & 2SLS & 2SLS \\
 & (LPM) & 2-way & 2-way & & & & (trends) \\
\midrule
GDP growth        & $-0.507^{**}$ & $-0.460^{**}$ & $-0.443^{**}$ &        &              & $0.141$       & $-1.297$ \\
                  & $(0.222)$     & $(0.220)$     & $(0.217)$     &        &              & $(1.330)$     & $(1.071)$ \\
GDP growth (lag)  & $0.020$       & $-0.173$      & $-0.121$      &        &              & $-0.165$      & $0.040$ \\
                  & $(0.243)$     & $(0.221)$     & $(0.277)$     &        &              & $(0.229)$     & $(0.185)$ \\
Rainfall growth   &               &               &               & $0.024$ & $0.042^{**}$ &              &  \\
                  &               &               &               & $(0.055)$ & $(0.018)$  &              &  \\
Rainfall growth (lag) &           &               &               & $-0.084^{*}$ & $0.025$ &              &  \\
                  &               &               &               & $(0.048)$ & $(0.016)$  &              &  \\
Polity~2 (lag)    & $0.000$       & $-0.004$      & $-0.002$      & $-0.003$ & $-0.001^{*}$ & $-0.003$    & $-0.004$ \\
                  & $(0.004)$     & $(0.005)$     & $(0.005)$     & $(0.005)$ & $(0.001)$   & $(0.006)$   & $(0.005)$ \\
\midrule
$N$               & 702 & 702 & 702 & 702 & 702 & 702 & 702 \\
$R^2$             & 0.129 & 0.012 & 0.006 & 0.005 & 0.019 & 0.569 & 0.710 \\
\bottomrule
\end{tabular}\\[2pt]
\footnotesize $^{*}p<0.1$, $^{**}p<0.05$, $^{***}p<0.01$.
\end{table}

POLS reproduces the MSS direction: a one-standard-deviation drop in
growth ($\approx 6.6$ percentage points) raises the conflict
probability by $0.51 \times 0.066 \approx 3.4$ percentage points,
significant at the 5\% level. FE and RE shrink the coefficient slightly
towards zero ($\hat\beta = -0.46, -0.44$) with widened standard errors,
as expected when between-country variation is stripped out. The reduced
form shows that rainfall growth enters the conflict equation only
weakly (lagged rainfall has $p = 0.08$). The first stage confirms that
current-year rainfall growth has a statistically significant but
quantitatively modest effect on GDP growth.

The 2SLS estimate (column 6) is the striking column. The point estimate
is $+0.14$ with a cluster-robust SE of $1.33$: the wrong sign, no
statistical significance, and a confidence interval covering essentially
all economically relevant magnitudes. This collapse of the IV estimate
relative to the well-determined POLS estimate is the canonical
fingerprint of a weak instrument. Column~7 re-estimates the same model
with country FE plus country-specific linear time trends (the MSS
specification) and recovers $\hat\beta = -1.30$ (SE $1.07$, $p = 0.23$):
the correct sign, larger in magnitude than OLS, but still imprecisely
estimated. The qualitative direction of MSS is therefore sensitive to
whether secular country-level trends are modelled.

\begin{table}[H]
\centering
\caption{Specification and identification diagnostics.}
\label{tab:diag}
\small
\begin{tabular}{lrr}
\toprule
Statistic & Value & $p$-value \\
\midrule
\textit{Heteroscedasticity} & & \\
Breusch--Pagan                                       & $142.52$ & $<10^{-16}$ \\
\textit{Panel structure} & & \\
$F$-test country FE ($H_0\colon c_i=0$)              & $14.90$  & $<10^{-16}$ \\
Hausman FE vs RE                                     & $0.04$   & $0.998$ \\
\textit{Weak instruments} & & \\
First-stage $F$ baseline (cluster-robust)            & $2.89$   & $0.056$ \\
First-stage $F$ MSS-trends (cluster Wald)            & $4.45$   & $0.012$ \\
First-stage $F$ MSS-trends (homoscedastic)           & $7.48$   & $0.0006$ \\
Stock--Yogo 10\%-maximum-size critical value         & $19.93$  & --- \\
\textit{Endogeneity} & & \\
Wu--Hausman                                          & $0.14$   & $0.705$ \\
\textit{Overidentification} & & \\
Sargan                                               & $2.40$   & $0.121$ \\
\textit{Bootstrap inference} & & \\
Cluster-bootstrap 95\% CI for $\beta^{\text{IV}}$    & $[-3.52,\,3.51]$ & --- \\
\bottomrule
\end{tabular}
\end{table}

The diagnostics tell a coherent story. Heteroscedasticity is decisively
present, confirming the necessity of robust standard errors for the LPM.
Country fixed effects are jointly significant, but the Hausman test
returns a vanishing statistic that we discount: the test's construction
assumes the FE estimator is efficient under $H_0$, which fails once
standard errors are clustered, and its power is limited in moderately
unbalanced panels. We default to FE.

The weak-instrument diagnostics are the crux. The first-stage $F$ on
the two excluded rainfall instruments is $2.89$ in our baseline,
roughly one-seventh of the Stock--Yogo threshold of $19.93$. The
MSS-trends specification raises it to $4.45$ (cluster-robust) or $7.48$
(homoscedastic), notably stronger but still below the threshold.
Conventional 2SLS inference is therefore unreliable in both
specifications, and the Wu--Hausman test fails to reject exogeneity
($p = 0.705$): in this sample we have no positive statistical evidence
that OLS is biased, though theory says it must be.

A non-parametric cluster bootstrap (500 replications, resampling
countries with replacement) yields a 95\% percentile interval of
$[-3.52,\,3.51]$, wide enough to include zero and most economically
relevant magnitudes. The Sargan test on the two rainfall instruments
fails to reject ($p = 0.121$), which we read as weak reassurance
only: with one degree of freedom and two instruments that share their
exogeneity argument, the test cannot detect homogeneous invalidity of
the kind \citet{Sarsons2015} flags.

\begin{table}[H]
\centering
\caption{Anderson-Hsiao first-difference IV. The structural
equation~\eqref{eq:structural} is first-differenced to remove country
fixed effects; $\Delta g^{y}_{it}$ is instrumented by the lagged level
$g^{y}_{i,t-1}$. Cluster-robust SE at country level in parentheses.}
\label{tab:ah}
\small
\begin{tabular}{lr}
\toprule
 & Anderson-Hsiao FD-IV \\
\midrule
GDP growth ($\Delta$)         & $-0.349$ \\
                              & $(0.383)$ \\
Polity~2 (lag, $\Delta$)      & $-0.005$ \\
                              & $(0.005)$ \\
\midrule
$N$ (country-years)           & 620 \\
Countries                     & 41 \\
First-stage $F$ on $g^{y}_{i,t-1}$ (cluster-robust) & 451.7 \\
\bottomrule
\end{tabular}
\end{table}

The Anderson-Hsiao estimate is negative but imprecise: $\hat\beta_{AH}
= -0.349$ with cluster-robust standard error $0.383$ ($p = 0.36$). The
first-stage $F$ on the lagged level $g^{y}_{i,t-1}$ is $451.7$, far
above the Stock-Yogo 10\%-maximum-size critical value of $16.4$ for one
instrument and one endogenous regressor. The strength of the
first stage reflects mean reversion in the country-level growth
series: high past growth predicts a large negative change in current
growth almost mechanically, which makes the lagged level a relevant
instrument by construction. Identification, however, rests on a
different assumption from MSS: past growth must be uncorrelated with
the change in unobserved conflict shocks,
$\mathrm{Cov}(g^{y}_{i,t-1}, \Delta u_{it}) = 0$. This is logically
independent from the rainfall exclusion restriction; the two
strategies fail or succeed for different reasons. The MSS sign appears
again, but the panel-internal design pins down only the direction.

% ===========================================================================
\section{Discussion}\label{sec:discussion}
% ===========================================================================
The empirical findings are mixed and the appropriate reading depends
on what one is willing to assume about identification.

OLS recovers the qualitative MSS direction with a tight standard
error: $\hat\beta_{\text{POLS}} = -0.507$. Under exogeneity of growth
(which the three endogeneity sources of Section~\ref{sec:intro} rule
out), this would be the most precise summary of the conditional
relationship between growth and conflict. As a causal estimate, it is
not defensible.

Under the MSS rainfall identification, conventional inference is
unreliable. The first-stage $F$ of $2.89$ in our two-way FE
specification is roughly one-seventh of the Stock-Yogo threshold; the
2SLS point estimate carries the wrong sign and a confidence interval
that covers the entire interesting parameter space. Restoring the MSS
country-trend specification raises the first stage to $4.45$ and
recovers $\hat\beta = -1.30$, of the correct sign but still
imprecisely estimated. The cluster bootstrap percentile interval
$[-3.52,\,3.51]$ confirms the order of magnitude of the irreducible
uncertainty in this design under conventional 2SLS, though the
bootstrap addresses finite-sample clustered inference and does not
itself restore identification under weak instruments.

Under the panel-internal Anderson-Hsiao identification, the negative
sign reappears ($\hat\beta_{\text{AH}} = -0.349$) but the estimate
remains imprecise. The first-stage strength is mechanical: growth's
mean reversion makes the lagged level a strong predictor of the first
difference, but this does not introduce new identifying variation
that is logically independent of the rainfall design's failures.

What the data support: the qualitative sign of the MSS effect is
robust across credible specifications under at least two distinct
identifying assumptions (rainfall exclusion, past-growth
orthogonality). What the data do not support: a precise magnitude, or
a credible defence of either identifying assumption against the
\citet{Sarsons2015} critique, which targets the rainfall channel
directly. The Sargan over-identification test on the two rainfall
instruments does not reject, but with one degree of freedom and two
instruments that share their exogeneity argument the test has neither
the power nor the structure to detect the kind of violation that
matters.

The most honest reading is that the data inform the sign of $\beta$
but not its magnitude. A precise causal estimate of how much income
shocks raise the probability of civil conflict in this sample requires
either a stronger instrument, the commodity-price approach of
\citet{BrucknerCiccone2010} being the leading candidate, or a research
design that does not rely on rainfall at all.

% ---------------------- bibliography ----------------------
\renewcommand{\refname}{References}
\begin{thebibliography}{99}\small

\bibitem[Anderson and Hsiao(1981)]{AndersonHsiao1981}
Anderson, T.W., and C.\ Hsiao (1981). ``Estimation of dynamic models
  with error components.'' \emph{Journal of the American Statistical
  Association} 76(375): 598--606.

\bibitem[Br\"uckner and Ciccone(2010)]{BrucknerCiccone2010}
Br\"uckner, M., and A.\ Ciccone (2010). ``International commodity
  prices, growth and the outbreak of civil war in Sub-Saharan Africa.''
  \emph{Economic Journal} 120(544): 519--534.

\bibitem[Ciccone(2011)]{Ciccone2011}
Ciccone, A.\ (2011). ``Economic shocks and civil conflict: A
  comment.'' \emph{American Economic Journal: Applied Economics}
  3(4): 215--227.

\bibitem[Grossman(1991)]{Grossman1991}
Grossman, H.I.\ (1991). ``A general equilibrium model of
  insurrections.'' \emph{American Economic Review} 81(4): 912--921.

\bibitem[Miguel et al.(2004)]{MSS2004}
Miguel, E., S.\ Satyanath, and E.\ Sergenti (2004). ``Economic
  shocks and civil conflict: An instrumental variables approach.''
  \emph{Journal of Political Economy} 112(4): 725--753.

\bibitem[Miguel and Satyanath(2011)]{MiguelSatyanath2011}
Miguel, E., and S.\ Satyanath (2011). ``Re-examining economic shocks
  and civil conflict.'' \emph{American Economic Journal: Applied
  Economics} 3(4): 228--232.

\bibitem[Sarsons(2015)]{Sarsons2015}
Sarsons, H.\ (2015). ``Rainfall and conflict: A cautionary tale.''
  \emph{Journal of Development Economics} 115: 62--72.

\bibitem[Staiger and Stock(1997)]{StaigerStock1997}
Staiger, D., and J.H.\ Stock (1997). ``Instrumental variables
  regression with weak instruments.'' \emph{Econometrica} 65(3):
  557--586.

\bibitem[Stock and Yogo(2005)]{StockYogo2005}
Stock, J.H., and M.\ Yogo (2005). ``Testing for weak instruments in
  linear IV regression.'' In: \emph{Identification and Inference for
  Econometric Models: Essays in Honor of Thomas Rothenberg}, ed.\
  D.W.K.\ Andrews and J.H.\ Stock, Cambridge University Press,
  80--108.

\end{thebibliography}

\end{document}
